Palaeodistribution of crop progenitors and wild foods in the Southern Levant

### 2023-08-17 analysis/analysis.Rmd
Tidy fitting

```{r dependencies}
library("dplyr")
library("fs")
library("furrr")
library("ggplot2")
library("ggspatial")
library("gt")
library("here")
library("parsnip")
library("purrr")
library("ranger")
library("readr")
library("rgbif")
library("rpaleoclim")
library("sf")
library("stars")
library("stringr")
library("tibble")
library("tidyr")

@@

```{r read-archaeo_flora}
archae_flora <- read_tsv(raw_data("swasia_neolithic_flora.tsv"))
```

@@

```{r filter-flora}
flora <- filter(flora, n_occ > 50)
```

@@

flora |>
  mutate(
    occ = future_map(taxon, gbif_data, 
                     hasCoordinate = TRUE, hasGeospatialIssue = FALSE, 
                     geometry = w_asia, limit = 100000)
  ) ->
  flora
```

```{r tidy-occ}
# TODO: further tidying. Ideas:
# - https://data-blog.gbif.org/post/gbif-filtering-guide/
# - https://cran.r-project.org/web/packages/CoordinateCleaner/vignettes/Cleaning_GBIF_data_with_CoordinateCleaner.html
# - https://www.r-bloggers.com/2021/03/downloading-and-cleaning-gbif-data-with-r/
flora |>
  mutate(
    occ = future_map(occ, gbif_drop_imprecise_coords, threshold = 5000),
    occ = future_map(occ, gbif_drop_fossils),
    occ = future_map(occ, gbif_drop_duplicate_coords),
    occ = future_map(occ, gbif_standardise_names),
    occ = future_map(occ, \(x) {
      select(x, gbif_key = key, scientific_name, genus, species, longitude, 
             latitude)
    })
  ) ->
  flora
```

@@

```{r generate-background}
# TODO: Refactor to work with nested occ
# TODO: maybe defer this to just before the model is actually fit 
#   (and discard the background sample)
n_background <- 10000

occ |>
  mutate(present = TRUE) |>
  group_by(taxon) |>
  nest(.key = "occ_data") |>
  mutate(bg_data = map(1:n(), \(i) { 
    background_sample(
      region = w_asia_land, 
      N = n_background,
      coord_x = "longitude", coord_y = "latitude",
      present = FALSE
    )
  })) |>
  mutate(data = map2(occ_data, bg_data, bind_rows)) |>
  select(taxon, data) |>
  unnest(data) |>
  st_sf() ->
  occ
```

@@

```{r training-climate}
occ <- cbind(occ, st_extract(climate$cur, occ)) |> select(-geometry.1)
```

@@

# Drop any rows with NA values in any of the predictor variables
# TODO: Where do these actually come from? There's usually only a handful.
occ <- drop_na(occ, rf_predictors())

@@

# Fit to each taxon
# TODO: parallelise
occ |>
  nest(.by = taxon) |>
  mutate(
    data = map(data, as_tibble),
    fit = map(data, \(data) {
      fit(rf_model, factor(present) ~ ., data = select(data, rf_pred_vars()))
    })) ->
  rf_fits

# TODO: not this
rf_fits$taxon <- taxa$taxon
```

>>>>

```{r dependencies}
library("dplyr")
library("fs")
library("furrr")
library("ggplot2")
library("ggspatial")
library("gt")
library("here")
library("parsnip")
library("progressr")
library("purrr")
library("ranger")
library("readr")
library("rgbif")
library("rpaleoclim")
library("sf")
library("stars")
library("stringr")
library("tibble")
library("tidyr")

@@

```{r read-archaeo_flora}
archaeo_flora <- read_tsv(raw_data("swasia_neolithic_flora.tsv"))
```

@@

```{r filter-flora}
flora <- filter(flora, n_occ > 50)
# TODO: Temporary measure.
# Why are there 47k occurrences of Avena sterilis??
flora <- filter(flora, taxon != "Avena sterilis")
```

@@

flora |>
  mutate(
    occ = map(taxon, gbif_data, 
              hasCoordinate = TRUE, hasGeospatialIssue = FALSE, 
              geometry = w_asia, limit = 100000,
              .progress = TRUE)
  ) ->
  flora
```

```{r tidy-occ}
# TODO: further tidying. Ideas:
# - https://data-blog.gbif.org/post/gbif-filtering-guide/
# - https://cran.r-project.org/web/packages/CoordinateCleaner/vignettes/Cleaning_GBIF_data_with_CoordinateCleaner.html
# - https://www.r-bloggers.com/2021/03/downloading-and-cleaning-gbif-data-with-r/
flora |>
  mutate(
    occ = map(occ, gbif_drop_imprecise_coords, threshold = 5000),
    occ = map(occ, gbif_drop_fossils),
    occ = map(occ, gbif_drop_duplicate_coords),
    occ = map(occ, gbif_standardise_names),
    occ = map(occ, \(x) {
      select(x, gbif_key = key, scientific_name, genus, species, longitude, 
             latitude)
    })
  ) ->
  flora
```

@@

```{r generate-background}
# TODO: maybe defer this to just before the model is actually fit ?
#   (and discard the background sample)
n_background <- 10000

flora |>
  mutate(
    occ = map(occ, mutate, present = TRUE),
    bg_data = map(1:n(), \(i) { 
      background_sample(
        region = w_asia_land, 
        N = n_background,
        coord_x = "longitude", coord_y = "latitude",
        present = FALSE
      )
    }),
    occ = map2(occ, bg_data, bind_rows)
  ) |>
  select(-bg_data) ->
  flora
```

@@

```{r training-climate}
flora |> 
  mutate(
    occ = map(occ, \(x) cbind(x, st_extract(climate$cur, gbif_to_sf(x))))
  ) ->
  flora
```

@@

# Drop any rows with NA values in any of the predictor variables
# TODO: Where do these actually come from? There's usually only a handful.
# occ <- drop_na(occ, rf_predictors())

@@

# Fit to each taxon
# TODO: parallelise
flora |>
  mutate(
    fit = map(occ, \(occ) {
      fit(rf_model, factor(present) ~ ., data = select(occ, rf_predictors()))
    }, .progress = TRUE)
  ) ->
  flora
```
